---
title: "wfkelly2_Mod4_Peer_Review"
author: "William Kelly"
date: "6/28/2021"
output: html_document
---

# Load Libraries Needed
```{r}
library(tidyverse)
library(lubridate)
```


# Read in Module 4 Peer Review Data
```{r}
df <- read.csv("mod4_peer_rev_data.csv")
summary(df)
```

# Update Column Datatypes
```{r}
df$Department <- as.factor(df$Department)
df$Category <- as.factor(df$Category)
df$Date <- mdy(df$Date)

typeof(df$CustomerCode) # Character Type Check

```

# Summary of Quantity and Price Columns
```{r}
summary(df$Quantity)
summary(df$Price)
```
As shown by the summary we can see that 
for Quantity Max=24, Min=1, Mean=11.31 NA's=7
for Price Max=50, Min=3, Mean=22.81 NA's=10


# Display Count of NA's in Quantity and Price Columns
```{r}
sum(is.na(df$Quantity))
sum(is.na(df$Price))
```

# Bar Chart - Category Column
```{r}
ggplot(data = df) +
  geom_bar(mapping = aes(x = Category, fill=Category)) +
  theme(axis.text = element_text(size=5)) +
  xlab("Category") + ylab("Count Number") + ggtitle("Bar Chart of Category")
```

# Bar Chart - Departments and their Revenue
```{r}
df$Revenue <- df$Price * df$Quantity

df %>% ggplot(aes(x= fct_reorder(Department,-Revenue), y=Revenue)) + 
  geom_bar(stat="identity", color='skyblue',fill='steelblue') +
  labs(title = 'Bar Plot-Revenue by Departments')
```

# Histogram (Price & Quantity)
```{r}
ggplot(df, aes(x=Price)) + 
  geom_histogram(binwidth = 0.5) +
  coord_cartesian(xlim = c(0, 35)) +
  labs(title = 'Histogram of Price')

ggplot(df, aes(x=Quantity)) + 
  geom_histogram(binwidth = 0.5) +
  labs(title = 'Histogram of Quantity')
```

# Box and Whisker Plots (Price & Quantity)
```{r}
ggplot(df, aes(x=Price)) + 
  geom_boxplot() + 
  coord_cartesian(xlim = c(0, 35)) +
  labs(title = 'Box Plot-Price') +
  xlab("Price") + ylab("Count")

ggplot(df, aes(x=Quantity)) + 
  geom_boxplot() + 
  labs(title = 'Box Plot-Quantity') +
  xlab("Quantity") + ylab("Count")
```

# Essay Strengths and Weaknesses PowerBI, Alteryx and R

In terms of PowerBI, the tool does a great job simplifying the process of creating visualizations with the point and click interface. Also is good at sharing results with others.

In terms of Alteryx, the tool does a good job analyzing the data with a simply UI but doesn't do as good of a job at creating Visualizations compared to PowerBI and R Studio. I really like how you can see each step of the process for analyzing data in graphical UI form.

In terms of R Studio does a great job at providing built in libraries like Tidyverse to simplify and provide added abilities to create charts and analyze the data. In summary I feel like R studio is the best in terms of ability, sharing, and replicability. 






