---
title: "wfkelly2_HW3"
author: "William Kelly"
date: "6/21/2021"
output: html_document
---

# Load Libraries
```{r}
library(tidyverse)
library(corrplot)
library(Hmisc)
```


# 1) Import store_3HE dataframe as df 
```{r}
df <- read.csv("store_3HE.csv")
```

# 2) Familiarize w/ the features of the data frame
```{r}
str(df)
summary(df)
head(df)
tail(df)
slice_sample(df)
# unique(df) 
n_distinct(df)
```
From the str function can see that there are 771 rows of data with 20 different columns.
Noticed from Summary function the revenue column is in character format.
Also can see that there are 15 NA's in the size column and 1 NA in Revenue col.

# 3) Convert Values in Revenue to numeric data type
```{r}
df$revenue <- as.numeric(df$revenue)
```

# 4) Delete rows that contain missing values in any column
```{r}
df <- df %>% filter(size != "NA", revenue != "NA")
```
Noticed 15 NAs in size column and 1 in revenue so filtered out rows w/ NA's (755 rows)

# 5) Create two dataframes df_low and df_high
```{r}
# Calculate mean of revenue from df
df_mean_revenue <- mean(df$revenue)
df_low <- df %>% filter(df$revenue < df_mean_revenue)
df_high <- df %>% filter(df$revenue >= df_mean_revenue)

# Average Size of each low/high
size_df_low <- mean(df_low$size)
size_df_high <- mean(df_high$size)


# Line plots for df_low and df_high showing Revenue by Size
ggplot(df_low,aes(x=revenue,y=size)) +
  geom_line() + 
  labs(title='Line Plot-Size by Revenue df_low')

ggplot(df_high,aes(x=revenue,y=size)) +
  geom_line() + 
  labs(title='Line Plot-Size by Revenue df_high')
```

One can see by the Line Plot that for both df_low and df_high the
Revenue and Size is a positive relationship, 
as Revenue goes up the size also increases.

# 6) Correlation between Revenue and Size of df
```{r}
cor(df$revenue,df$size)

ggplot(df,aes(x=revenue,y=size)) +
  geom_line() + 
  labs(title='Line Plot-Size by Revenue df')
```

As shown in the line plot for df dataframe the relationship between
Revenue and Size is a positive slope meaning as Revenue increases, Size increases.
The Correlation calculated for Revenue and Size is 0.6175055.

# 7) Boxplots of Revenue for each region
```{r}
ggplot(df, aes(x=region, y=revenue)) + 
  geom_boxplot() +
  labs(title = 'Box Plot-Revenue by Region')
```

As shown in the boxplots for each region, 
the average revenue is greater in the West & Quebec compared to Atlantic & Ontario.
Also there are more higher values in the West noted by larger third quartile from mean.
In addition, Ontario has the lowest mean and spread of revenue.

# 8) Correlation between revenue and promo_units
```{r}
cor(df$revenue,df$promo_units)

ggplot(df,aes(x=revenue,y=promo_units)) +
  geom_line() + 
  labs(title='Line Plot-Promo_units by Revenue df')

```

Correlation between revenue and promo_units is 0.9659697
This aligns with expectation in that more promo units result in higher revenue.
Based on the Line Plot trend I believe that selling more units on promotion 
will increase revenue due to the positive slope/correlation b/t the two variables.

# 9) Bar Chart showing total gross_profit for each province
```{r}
ggplot(df, aes(x=province, y=gross_profit)) + 
  geom_col() + 
  labs(title = 'Bar Plot-Gross Profit by Province')
```

From the Bar Plot we can see that the provinces ON, AB and QC have the highest gross profit.
While the provinces NB, PE, and Unknown are the lowest in gross profit.

# 10) Correlation Matrix b/t Average Store Sales for all Product Categories
```{r}
Product_Categories <- df[,c(10,11,12,13,14,15,16,17,18,19,20)]
head(Product_Categories)
	
Prod_Cat.cor = cor(Product_Categories, method = c("spearman"))
corrplot(as.matrix(Prod_Cat.cor))

palette = colorRampPalette(c("green", "white", "red")) (20)
heatmap(x = Prod_Cat.cor, col = palette, symm = TRUE)
```

From the Correlation Matrix the product categories 
that sell the least together are psd591MI_units & gum_units.
While product categories bagpegCandy_units and takeHomePotato sell the most together.
