************************************************************************
# Title: Module 3HE ETL and EDA
Course: MBA563
Term: Summer 2021
Mooc or HE?: HE
Module: 03
Author: Jessen Hobson
************************************************************************

*********ETL & EDA********************************************
In class notebook
*********ETL & EDA********************************************

# Install and load packages (don't install twice)
```{r}
#install.packages('tidyverse')
library(tidyverse)
```

# 1.1 - Bring in the data and visualize
# *****Data unavailable*****
```{r}
df1 <- read_rds("Mod3HE_HMPV_eda.rds")
```

```{r}
summary(df1)
```

```{r}
str(df1)
```

```{r}
head(df1, n=5) #Multiple observations per participant. Lots of features.
```

```{r}
n_distinct(df1$jobtitle)
```

```{r}
unique(df1$jobtitle)
```


# 1.2 - Clean the data with the goal in mind



# QUESTION: What features do we care about?



# Subset the data (order here also determines the order of the variables)
```{r}
df2 <- df1 %>%
  select(subject, company, decep_firm, decep_judg, aud_exp, cdmanipulation, 
         overall_time, jobtitle, auditor_fraud, confidence, flagchars)
```

```{r}
# Examine and describe each feature
slice_sample(df2, n=10)
```

```{r}
df2 <- read_rds("Mod3HE_HMPV_eda.rds")
summary(df2)
```



# QUESTION: What do we do about NAs in `auditor_fraud`


# Rename a variable
```{r}
df2 <- df2 %>% 
  rename(prompt = cdmanipulation,
         judgment = decep_judg,
         fraud = decep_firm)
```


# Create variables using `mutate()`, `if_else()`, and `case_when()`
```{r}
df2 <- df2 %>% 
  mutate(accurate = if_else(fraud==judgment, 1, 0))

head(df2, n=3)
```



```{r}
df2 <- df2 %>% mutate(fraud = if_else(fraud==1, 'fraud firm', "clean firm"))

df2 <- df2 %>% mutate(judgment = if_else(judgment==1, 'say fraud', "say clean"))

df2 <- df2 %>% mutate(prompt = if_else(prompt==1, 'prompted', "unprompted"))

head(df2, n=3)
```



# Create professional dummy variable

# QUESTION: How do we know whether someone is an auditor or not? What variables should we look at? See PowerPoint.



```{r}
ggplot(df2, aes(x=aud_exp)) + 
  geom_histogram(binwidth = 1) +
  theme_minimal() +
  labs(title = 'Histogram of Audit Experience')
```



```{r}
df2 %>% dplyr::filter(aud_exp==1) %>% 
  select(aud_exp, jobtitle)
```


```{r}
df2 <- df2 %>% 
  mutate(professional = 
           case_when(aud_exp < 1 ~ 'student',
                     aud_exp > 1 ~ 'pro', 
                     aud_exp==1 & jobtitle == 'student' ~ 'student',
                     aud_exp==1 & jobtitle != 'student' ~ 'pro'))

df2 %>% filter(aud_exp==1) %>% 
  select(aud_exp, jobtitle, professional)
```


# 2.1 - Outliers


# QUESTION: Look at the summary. Do any features have potential outliers?
```{r}
summary(df2)
```



```{r}
ggplot(df2, aes(y=overall_time)) + 
  geom_boxplot() +
  theme_minimal()
```

```{r}
ggplot(df2, aes(y=overall_time)) + 
  geom_boxplot() + 
  coord_cartesian(ylim=c(0, 10000)) + 
  theme_minimal()
```

```{r}
df2 %>% 
  select(subject, overall_time, professional) %>%
  filter(overall_time > 3*sd(df2$overall_time) | overall_time < -3*sd(df2$overall_time)) %>% 
  mutate(minutes = overall_time/60, hours = overall_time/(60*60), days = overall_time/(60*60*24)) %>% 
  distinct()

```

```{r}
df2 <- df2 %>% filter(!subject %in% c(81, 196, 209))
summary(df2$overall_time)
```



# QUESTION: In groups, do a simple visualization of the data. Pick one or two variables to look at and create a plot with `ggplot()`. 

# Use data “Mod3HE_HMPV_eda.rds”
```{r}
df2 <- read_rds('Mod3HE_HMPV_eda.rds')
summary(df2)
```



# 2.2a Correlations
```{r}
cor(df2$confidence, df2$aud_exp)
```

```{r}
df2 %>% 
  select(confidence, flagchars, overall_time, aud_exp) %>% 
  cor()
```



# 2.2b Main Effects

# `group_by` and `summarize()`
```{r}
main_effects <- df2 %>% 
  group_by(professional) %>%
  summarize(accuracy = mean(accurate))

print(main_effects)
```

# Plot
```{r}
main_effects %>% 
  ggplot(aes(x=professional, y=accuracy, fill=professional)) +
  geom_col()
```

```{r}
main_effects %>% 
  ggplot(aes(x=professional, y=accuracy, fill=professional)) +
  geom_text(aes(label=round(accuracy, digits = 2)), vjust = -0.2) +
  geom_col()
```



# By type of firm and professional
```{r}
main_effects2 <- df2 %>% 
  group_by(professional, fraud) %>%
  summarize(accuracy = mean(accurate))

print(main_effects2)
```

```{r}
main_effects2 %>% 
  ggplot(aes(x=fraud, y=accuracy, fill=professional)) +
  geom_col(position='dodge2') + 
  geom_text(aes(label=round(accuracy, digits = 2)), vjust = -0.2, position = position_dodge(.9)) +
  theme_minimal()
```



# 2.3 - Interaction
```{r}
df2 %>% group_by(professional, fraud, prompt) %>% 
  summarize(accuracy = mean(accurate)) %>% 
  ggplot(aes(x=fraud, y=accuracy, fill=prompt)) +
  geom_col(position='dodge2') + 
  facet_grid(rows = professional~.) +
  geom_text(aes(label=round(accuracy, digits = 2)), vjust = 1.2, position = position_dodge(.9)) +
  theme_classic()
```