---
title: "mod1_peer_review_Bike_Share"
author: "William Kelly"
date: "7/7/2021"
output: html_document
---

# Load Libraries
```{r}
library(tidyverse)
library(lubridate)
library(GGally)
```

# Read in Bike Sharing Dataset "day.csv"
```{r}
df <- read.csv("day.csv")
```

# Structure and Summary of Data
```{r}
str(df)
summary(df)
```

# Convert dteday to month_name column
```{r}
df$dteday <- ymd(df$dteday)
df$month_name <- as.character(month(df$dteday, label = TRUE))
```

# Simple Linear Regression cnt (dep) & month_name (indep)
```{r}
model1 <- lm(cnt ~ month_name, data = df)
summary(model1)
```

The R Squared value is 0.3906 or 39.06% of the variation in cnt explained by month_name.
And the adjusted R Squared is 38.1% which shows that the fitted values are below the actual values of 39.06%. This shows a positive correlation & also show less variation in the regression model meaning less variation in dependent variable explained by indep variable.
Also the P-Value is small showing more statistical confidence in the model.

# GGpairs Plot for month_name and cnt
```{r}
ggpairs(df
        , columns = c(16,17) # Only include month_name and cnt
        , aes(color = month_name # Let's see if we can identify any difference among campaign. 
              , alpha = .4 # Makes the colors partially transparent
              )
        )
```


# Predict which month set as reference and cnt (April is Reference)
```{r}
ggplot(model1, aes(x = month_name, y = cnt)) +
  geom_point() +
  expand_limits(x = c(0,1)) +
  stat_smooth(method = 'lm', fullrange = T)

newObservations <- data.frame(month_name = 'Apr'
                              , cnt = c(2000, 3000, 4000) )

newObservations$predictedCnt = predict(model1, newObservations)

```

As shown in the summary of Model1, April is set as the reference month.
The predicted count for April would be, cnt = 4484.9


# Use Coefficients to estimate cnt for months of January and June
```{r}
cnt_Jan <- 4484.9 + -2308.6
cnt_Jan
cnt_Jun <- 4484.9 + 1287.5
cnt_Jun
```

# Multiple Linear Regression cnt (dep) & temp/month_name (indep)
```{r}
model2 <- lm(cnt ~ temp + month_name, data = df)
```

# Summary of Model2
```{r}
summary(model2)
```

***Explain why the R-squared is different from the two simple regression models

Model2 is different from Model1 due to adding in the temp into the Linear Regression Model.
Since temp has a higher coefficient meaning higher correlation, the R squared value is a little bit higher than the R squared in Model1.

***Compare the coefficient estimates for the month_nameJan variable in Model1 and Model2. 

With regular text explain why the coefficient estimates are different.
Model1 Jan Coefficient is -2308.6 while Model2 Jan coeff is -852.31.
They are different because when adding in the Temp variable, the correlation increases which is shown by the increase in R squared causing a better correlation in Model2 Jan coeff. 


***With either a code chunk or regular text, use the coefficient estimates from Model2 to report the predicted cnt for the month of January when the temperature is .25
```{r}
cnt_Jan_Temp = 1554.39 + -852.31 + 0.25 * 6235.14
cnt_Jan_Temp
```

























# 
