MBA 561 Introduction to Business Analytics with R
---
title: "mod3_peer_review_William_Kelly"
author: "William Kelly"
date: "4/29/2021"
output: html_document
---

# Note: Function sys date presents today date and reformatted into m,d,y format
```{r}
# 3)
# Use a code chunk to create a datetime object d that stores today’s date. 
# Only use base R (don’t use lubridate package functions). 
# The date must be in the “mm/dd/yyyy” format.

d <- format(Sys.Date(), format="%m/%d/%Y")

```
# Note: Checks data type and prints today's date
```{r}
# 4)
# Use a new code chunk to check the data type of the d object by printing it.

class(d)
print(d)
```
# Note: Loads lubridate package
```{r}
# 5)
# Use a new code chunk to install and load the package “lubridate.”
#install.packages("lubridate") #(Commented out, already installed package)

library("lubridate")
```
# Note: Code chunk extracts year, month, week, day and assigns to respective variables
```{r}
#6)
# In a new code chunk, use the appropriate functions from the lubridate package to extract the year, month, week and day from the object d created above and save them as d_year, d_month, d_week, and d_day respectively.

d <- mdy(d) # Reformat date into UTC format
year(d) # Print Year
d_year <- year(d)
month(d, label = T) # Print Month
d_month <- month(d, label = T)
week(d) # Print number of week
d_week <- week(d)
wday(d, label = T) # Print day of week 
d_day <- wday(d, label = T)
```
# Note: Calculates 25 days from today's date
```{r}
#7)
# Use a new code chunk to create another datetime object, d_25, using the object d, 
# where d_25 is the date 25 days from now.

d_25 <- d + days(25) 

```
# Note: Uses difftime function to calculate number of days between d and d_25 
```{r}
#8)
# Finally, use a new code chunk to calculate the difference between d and d_25 
# using the difftime function and check whether the difference is 25 days.

difftime(d_25, d, units = 'days') 

```




