---
title: "HE Assignment 5 – Regression"
author: "W Kelly, Eric Boyle, Quaidzoher Dhilla, Brandon Hyer, Christian Mukania, Vinh Tran"
date: "7/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Libraries
```{r}
library(tidyverse)
library(magrittr)
library(lubridate)
library(corrplot)
```

# Read in hmdaInterestRate.rds file
```{r}
df <- read_rds("hmdaInterestRate.rds")
```

# Report Structure of dataframe
```{r}
str(df)
```


# Divide by 1,000 for columns loan_amount, property_value, and income
```{r}
df$loan_amount <- df$loan_amount / 1000
df$property_value <- df$property_value / 1000
df$income <- df$income / 1000
```


# Create new column "Itp" dividing loan_amount by property_value
```{r}
df$Itp <- df$loan_amount / df$property_value
```


# Filter Data to keep observations which income is less than 300 (i.e. $300,000)
```{r}
df <- df %>% filter(income < 300)
```


# Report Summary of all Columns
```{r}
summary(df)
```


# Corr plot for interest_rate, Itp, income, applicant_age, property_value, and loan_amount
```{r}
df_corr <- cor(df[,c('interest_rate', 'Itp', 'income','applicant_age','property_value','loan_amount')])
df_corr

corrplot(df_corr
         , method = 'color' # I also like pie and ellipse
         , order = 'hclust' # Orders the variables so that ones that behave similarly are placed next to each other
         , addCoef.col = 'black'
         , number.cex = .6 # Lower values decrease the size of the numbers in the cells
         )
```

***Below the plot, identify what variable has the strongest negative correlation with interest_rate. Comment on what might explain why that correlation is negative.

The Variable with the strongest negative correlation is "loan_amount" (-0.33)

The correlation b/t loan amount and interest rates are negative b/c the higher the loan amount the lower the interest rate due to the larger down payment because lenders see a lower level of risk when you have more at stake in the property compared to smaller loan amounts.


# Regress interest rate on Itp
```{r}
lm1 <- lm(interest_rate ~ Itp, data = df)
summary(lm1)
```
***Interpret the coefficient estimate on ltp.

Conclude from lm1 that the coefficient for Itp is -1.05188 which shows a negative coefficient and intercept of 5.31852. 
Also the R squared value is 0.1017 which shows that it is not significantly correlated.   

# Regress interest rate on Itp and loan_amount
```{r}
lm2 <- lm(interest_rate ~ Itp + loan_amount, data = df)
summary(lm2)
```
***Comment on the change in the adjusted R-squared, as well as the change in the coefficient on ltp.

The R Squared value in lm2 improved which is now 0.1522 but not significantly.
The coefficient for Itp is now -0.737 which shows lm2 has an increased correlation.
The coefficient for loan_amount is -0.002 which is a small correlation.


# Regress interest rate on Itp, loan_amount, and aus_1
```{r}
lm3 <- lm(interest_rate ~ Itp + loan_amount + aus_1, data = df)
summary(lm3)
```
***Interpret the new coefficients.

From lm3 we can see that the R squared value increased with stronger correlation (0.2675)
Also in aus_1 the value Desktop Underwriter (DU) is set as the reference.
For coefficients Itp is 0.025007, loan_amount is -0.001714 which is an increase in correlation from the previous models.
In addition from aus_1 there is a strong positive correlation for Not Applicable and Other.

